---
title: "zoo-clustering-kMeans"
author: "Bei Yu & Tianlin Liu"
date: "March 29, 2016"
output: word_document
---

This is a tutorial on using the kMeans algorithm in the RWeka package to cluster animal types in a zoo.

# load in data
```{r}
zoo <- read.csv("/Users/byu/Desktop/Data/zoo.csv")    
str(zoo)
```

# remove columns "names" and "type"
```{r}
zoo_unlabel <- zoo[, c(2:17)]
str(zoo_unlabel)
```

# build kMeans model using RWeka
```{r}
library(RWeka)
model_rweka <- SimpleKMeans (zoo_unlabel, control = Weka_control(N = 7, I=500, S=100))
model_rweka
```

# Since it is difficult to interpret and visualize the clustering results with RWeka, we will introduce some built-in R functions, as well as visualization packages

## use the kMeans algorith in R
```{r}
model_r <- kmeans(zoo_unlabel, 7)
model_r
```

## print the centroids
```{r}
model_r$centers
```

## get cluster assignment
```{r}
cluster_assignment <- data.frame(zoo,model_r$cluster)
View(cluster_assignment)
```

## visualize animal types and clusters by specific features, red = milk yes, black = milk no.
```{r}
plot(zoo$type ~ jitter(model_r$cluster, 1), pch=21,col=as.factor(zoo$milk)) 
```

## use PCA in visualization package "cluster" to visualize kMeans model. PCA is principal components analysis
```{r}
#install.packages("cluster")
library(cluster)
clusplot(zoo_unlabel,model_r$cluster,color=TRUE,shade=TRUE, labels=2, lines=0) # plot clusters
```

# use R's HAC algorithm, which uses Euclidean distance and complete linkage by default.
```{r}
d = dist(as.matrix(zoo_unlabel))
hc=hclust(d)
plot(hc)
```
