---
title: "groceries"
output: word_document
---

This is a tutorial uses R package "arules" to analyze the market basket data. 

# Load the libraries
```{r}
# install.packages(arules)
# install.packages(arulesViz)
library(arules)
library(arulesViz)
```

# Load the dataset
```{r}
groceries <- read.transactions("/Users/byu/Desktop/Data/groceries.csv", format="basket", sep="," )
```

# We can explore the data before we make any rules. The following script is going to create an item frequency plot for the top 20 items.
# "type" indicates whether item frequencies should be displayed relative or absolute
```{r}
itemFrequencyPlot(groceries,topN=20,type="absolute")
```

# Mine rules with the Association Rule algorithm. 
# It is required to set the minimum support and confidence values.
```{r}
rules <- apriori(groceries, parameter = list(supp = 0.001, conf = 0.8))
```

# Show the top 5 rules, rounding with 2 digits
# From the top 5 rules, we could notice that the rule {bottled beer, soups} => {whole milk} is strong, indicating bottled beer and soups are frequently bought together with whole milk. So we would suggest to place these items close to each other.
```{r}
options(digits=2)
inspect(rules[1:5])
```

# Get summary info about all rules
```{r}
summary(rules)
```

# Sort rules so that we can view the most relevant rules first. For example, sort rules with "confidence":
```{r}
rules<-sort(rules, by="confidence", decreasing=TRUE)
```

# If we want to target items to generate rules. For example, the frequently bought items with “whole milk”:
# "minlen" is to avoid empty left hand side items.
```{r}
rules<-apriori(data=groceries, parameter=list(supp=0.001,conf = 0.08, minlen=2), 
               appearance = list(default="lhs",rhs="whole milk"),
               control = list(verbose=F))
rules<-sort(rules, decreasing=TRUE,by='confidence')
inspect(rules[1:5])
```

# Likewise, we could set "whole milk" on the left hand side as well.
```{r}
rules<-apriori(data=groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), 
               appearance = list(default="rhs",lhs="whole milk"),
               control = list(verbose=F))
rules<-sort(rules, decreasing=TRUE,by="confidence")
inspect(rules[1:5])
```

# Visualize the rules
```{r}
#plot(rules,method="graph",interactive=TRUE,shading=NA)
```